---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Code for reading in the dataset and/or processing the data
```{r message=FALSE, warning=FALSE}
library(readr)
activity <- read_csv("activity.zip")
```
## What is mean total number of steps taken per day?
The total number of steps taken per day:
```{r message=FALSE, warning=FALSE}
library(tidyverse)
(total_steps <- activity %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps)))
```
2. Histogram of the total number of steps taken each day
```{r Histogram_of_the_total_number_of_steps_taken_each_day}
with(total_steps, hist(total_steps, main = "Histogram of total steps per day", xlab = "Total steps per day"))
```
```{r}
mean_steps <- mean(total_steps$total_steps, na.rm = T)
median_steps <- median(total_steps$total_steps, na.rm = T)
```
3. Mean and median number of steps taken each day  
The mean of the total number of steps taken per day is `r as.integer(mean_steps)` steps.  
The median of the total number of steps taken per day is `r median_steps` steps.  

## What is the average daily activity pattern?
4. Time series plot of the average number of steps taken
```{r Time_series_plot_of_the_average_number_of_steps_taken}
avg_steps <- activity %>%
  group_by(interval) %>%
  summarise(avg_steps = mean(steps, na.rm = T))

avg_steps %>%
  ggplot(aes(interval, avg_steps)) +
    geom_line()
```

5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
interval_with_highest_steps <- avg_steps %>%
  arrange(desc(avg_steps)) %>% .$interval %>% .[1]
```
The `r interval_with_highest_steps`th 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.  

## Imputing missing values
```{r}
# verify that NAs only occur in "steps" column
all(!is.na(activity[,2:3]))

# calculate the number of NAs
total_NA <- sum(is.na(activity$steps))
```
There are total `r total_NA` missing values in the dataset (i.e. the total number of rows with NA).
  
6. Code to describe and show a strategy for imputing missing data: impute the NAs with the mean for that 5-minute interval
```{r}
activity_NA_imputed <- activity
activity_NA_imputed[is.na(activity_NA_imputed$steps),] <- activity_NA_imputed %>%
  filter(is.na(steps)) %>%
  left_join(avg_steps, by = "interval") %>%
  mutate(steps = round(avg_steps)) %>% select(1:3)
activity_NA_imputed
```
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r Histogram_of_the_total_number_of_steps_taken_each_day_after_missing_values_are_imputed}
total_steps_NA_imputed <- activity_NA_imputed %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps))

with(total_steps_NA_imputed, hist(total_steps, main = "Histogram of total steps per day", xlab = "Total steps per day"))
```

```{r}
mean_steps_NA_imputed <- mean(total_steps_NA_imputed$total_steps, na.rm = T)
median_steps_NA_imputed <- median(total_steps_NA_imputed$total_steps, na.rm = T)
```
The mean of the total number of steps taken per day with NA-imputed data is `r as.integer(mean_steps_NA_imputed)` steps.  
The median of the total number of steps taken per day with NA-imputed data is `r as.integer(median_steps_NA_imputed)` steps.  
Both mean and median total number of steps taken per day are slightly smaller than previously calculated.  

## Are there differences in activity patterns between weekdays and weekends?
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r Panel_plot_comparing_the_average_number_of_steps_taken_per_5-minute_interval_across_weekdays_and_weekends, message=FALSE, warning=FALSE}
library(lubridate)

activity_NA_imputed %>% 
  mutate(weekday = as.factor(ifelse(wday(date, locale = "us")%in%2:6, "weekday","weekend"))) %>%
  group_by(weekday, interval) %>%
  summarise(avg_steps = mean(steps)) %>%
  ggplot(aes(interval, avg_steps)) +
  geom_line() + facet_wrap(~weekday, nrow = 2) +
  labs(y = "Number of steps")
```

